---
title: "BikeLM Lecture"
author: "NS"
date: "2023-03-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = ls())
set.seed(33)
```

```{r load packages}
library(tidyverse)
library(readr)
library(lubridate)
library(dplyr)
library(data.table)
library(leaps)
library(ggplot2)
library(dplyr)
library(MASS)
library(corrplot)
```

## Overview

**Objective:**

Create a linear model to estimate the demand for bikes and test its accuracy. 
Create transformations on the linear model and test their accuracy.

**Response Variables:**

*Y*: Hourly count rentals of bikes

**Qualitative Predicting Variables:**

*X1* = Day of the week
*X2* = Month of the year
*X3* = Hour of the day (ranging 0-23)
*X4* = Year (2011, 2012)
*X5* = Holiday Indicator
*X6* = Weather condition (with four levels from good weather for level 1 to severe condition for level 4)

**Quantitative Predicting Variables:**

*X7* = Normalized temperature
*X8* = Normalized humidity
*X9* = Wind speed

### Load Data

```{r load data}
data <- read.csv("Bikes.csv", header=T)
head(data)
```

## Preparing the Data

### As.Factor for Categorical

```{r preparing data}
# Set a seed for reproducibility
set.seed(9)

# Remove the irrelevant columns
clean_data = data[-c(1,2,9,15,16)]

# Convert the numerical categorical variables to predictors
clean_data$season = as.factor(clean_data$season)
clean_data$yr = as.factor(clean_data$yr)
clean_data$mnth = as.factor(clean_data$mnth)
clean_data$hr = as.factor(clean_data$hr)
clean_data$holiday = as.factor(clean_data$holiday)
clean_data$weekday = as.factor(clean_data$weekday)
clean_data$weathersit = as.factor(clean_data$weathersit)
```

## Regression Model

We fit a linear regression model with all the data
```{r regression model}
model1 = lm(cnt~., data=clean_data)
summary(model1)
```

### Find insignificant values

We want to identify p-values larger than sig level 0.05.
```{r insignificant values}
# Find insignificant values
which(summary(model1)$coeff[,4]>0.05)
```
## GOF Residual Analysis
```{r Constant Variance}
resids = rstandard(model1)
fits = model1$fitted
plot(fits, resids, xlab="Fitted Values", ylab="Residuals", main="Scatterplot", col="darkblue")
```
Does not hold. Seems to be constant then flares. Need some controlling
```{r Linearity Assumption}
par(mfrow = c(2,2))
plot(data$temp, resids, xlab = "Temperature", ylab = "Residuals")
abline(h = 0)
plot(data$atemp, resids, xlab = "Feeling Temperature", ylab = "Residuals")
abline(h = 0)
plot(data$hum, resids, xlab = "Humidity", ylab = "Residuals")
abline(h = 0)
plot(data$windspeed, resids, xlab = "Windspeed", ylab = "Residuals")
abline(h = 0)
```
All residuals are ok, no transformation needed
```{r Normality}
par(mfrow = c(1,2))
hist(resids, nclass=20, col="darkblue", border="gold", main="Histogram of residuals")

qqnorm(resids,col="darkblue")
qqline(resids, col="red")
```

### Goodness of Fit: outliers

```{r cooks dist and outliers}
# Cook’s Distance
cook = cooks.distance(model1)
plot(cook, type="h", lwd=3, col="darkred", ylab = "Cook's Distance", main="Cook's Distance")
```

It looks like theres an outlier, but it is so small at less than 0.004, so we don't use it

## Transformation of Response Variable with BoxCox
Assumption of constant variance doesn't hold, so we need to stabiliize the respones variable
We are going to use the Box-Cox
```{r BoxCox}
## Box Cox transformation
bc <- boxCox(model1)
lambda <- bc$x[which(bc$y==max(bc$y))]
lambda
```
Our lambda is 0.22, so we use 0.5 since it is count dat per unit time.

## Model with Transformation

```{r transformation}
## Fitting the model with square root transformation
model2<-lm(sqrt(cnt)~.,data=clean_data)
summary(model2)
```
The R^2 increased from without this trasformation

## Multicollinearity VIF
we look for multicollinearity. We look for max(10, 1/(1-R2)). If it's over, it's a problem

```{r multicollinearity VIF}
vif(model2)

```

There is a problem with season, month, temp, and atemp. This means we should not use all of the predictors in teh model.

## Removing Low Demand
We said before hours 0-6 has low demand so could disrupt the model

```{r remove low demand}
# Set a seed for reproducibility
set.seed(9)

# Remove data for hours 0-6
hrs = as.numeric(data$hr) #change hours to NUMERIC so we can do the greater than
data_red = clean_data[which(hrs>=7),] #only chose hours over 7

# Fitting the model with square root transformation
model3 = lm(sqrt(cnt)~.,data=data_red)
summary(model3)$r.squared

which(summary(model3)$coeff[,4]>0.05)
```
The R-squared is less, so maybe not so much better

The implication of the constant variation assumption violation is that the uncertainty in predicting bike demand when in high demand will be higher than estimated using the multiple regression model

## Prediction

### split to testing and training
```{r prediction training testing}
set.seed(9)

# Random Sampling Approach: 80% Train 20% Test split
sample_size = floor(0.8*nrow(clean_data))
picked = sample(seq_len(nrow(clean_data)), size=sample_size)
train = clean_data[picked,]
test = clean_data[-picked,]

sample_size_red = floor(0.8*nrow(data_red))
picked_red = sample(seq_len(nrow(data_red)), size=sample_size_red)
train_red = data_red[picked_red,]
test_red = data_red[-picked_red,]

##Apply the three models to training data
model1 =  lm(cnt ~ ., data=train)  #regular
model2 = lm(sqrt(cnt) ~ ., data=train) #transformed response
model3 = lm(sqrt(cnt) ~ ., data= train_red) #reduced predictors

# Build a prediction of the models with the new test data
# Specify whether a confidence or prediction interval
pred1 = predict(model1, test, interval = 'prediction')
pred2 = predict(model2, test, interval = 'prediction')
pred3 = predict(model3, test_red, interval = 'prediction')

head(pred1, 12)
```

The prediction has larg intervals, so maybe not so great.

##Prediction Accuracy with MSPE MAE etc

```{r accuracy}
## Save Predictions to compare with observed data
test.pred1 = pred1[,1]
test.lwr1 = pred1[,2]
test.upr1 = pred1[,3]

## Prediction measure functions
# Mean Squared Prediction Error (MSPE)
mse_fun <- function(pred,dat){mean((pred-dat)^2)}
mse_fun(test.pred1,test$cnt)

# Mean Absolute Prediction Error (MAE)
mae_fun <- function(pred,dat){mean(abs(pred-dat))}
mae_fun(test.pred1,test$cnt)

# Mean Absolute Percentage Error (MAPE)
mape_fun <- function(pred,dat){mean(abs(pred-dat)/abs(dat))}
mape_fun(test.pred1,test$cnt)

# Precision Measure (PM)
pm_fun <- function(pred,dat){sum((pred-dat)^2)/sum((dat-mean(dat))^2) }
pm_fun(test.pred1,test$cnt)
       
# CI Measure (CIM)
ci_fun <- function(pred.lwr,pred.upr,dat){
  sum(dat<pred.lwr|dat>pred.upr)/length(dat) 
}
ci_fun(test.lwr1,test.upr1, test$cnt)
```

MSE depends on magnitude of data, so its' meaningless here. Same with MAE.
MAP is 2.72, which is large, because avg diffrence between forecases value and actual is 272%
Precision is 0.31, meaning variable in the prediction is 31% variableity in new response. This is good since smaller than one.
CIM, we find that approximately 6% of the new response data is outside of the 95% prediction intervals.


## Prediction Accuracy: Averaging (Model 1)

This is a function
```{r prediction function}
## Prediction Function
pred_fun <- function(model,test,transf=FALSE){
 pred = predict(model, test, interval = 'prediction')
 test.pred = pred[,1]
 test.lwr = pred[,2]
 test.upr = pred[,3]
 if(transf==FALSE){
   mse_model = mse_fun(test.pred,test$cnt) 
   mae_model =  mae_fun(test.pred,test$cnt)
   mape_model = mape_fun(test.pred,test$cnt) 
   pm_model  = pm_fun(test.pred,test$cnt) 
   ci_model =  ci_fun(test.lwr,test.upr,test$cnt)
   predict_meas = c(mse_model,mae_model,mape_model,pm_model,ci_model)
 }
 else{
   mse_model = mse_fun(test.pred^2,test$cnt) 
   mae_model =  mae_fun(test.pred^2,test$cnt)
   mape_model = mape_fun(test.pred^2,test$cnt) 
   pm_model  = pm_fun(test.pred^2,test$cnt) 
   ci_model =  ci_fun(test.lwr^2,test.upr^2,test$cnt)
   predict_meas = c(mse_model,mae_model,mape_model,pm_model,ci_model)
 }
 return(predict_meas)
}
```


```{r sampling for 100 times}
## Apply Sampling & Predictions 100 times then average
pred1_meas = matrix(0,5,100)
pred2_meas = matrix(0,5,100)
pred3_meas = matrix(0,5,100)

## Apply Random Sampling 100 times
for(i in 1:100){
# Random Sampling Approach: 80% Training Data & 20% Testing Data
  sample_size = floor(0.8*nrow(clean_data))
  picked = sample(seq_len(nrow(clean_data)), size=sample_size)
  train = clean_data[picked,]
  test = clean_data[-picked,]
  
  sample_size_red = floor(0.8*nrow(data_red))
  picked_red = sample(seq_len(nrow(data_red)), size=sample_size_red)
  train_red = data_red[picked_red,]
  test_red = data_red[-picked_red,]
  
  # Apply the two models to training data 
  model1.train =  lm(cnt ~ ., data=train) 
  model2.train = lm(sqrt(cnt) ~ ., data=train) 
  model3.train = lm(sqrt(cnt) ~ ., data= train_red)
  
  ## Save Predictions for the i-th iteration
  pred1_meas[,i] = pred_fun(model1.train,test)
  pred2_meas[,i] = pred_fun(model2.train,test,TRUE)
  pred3_meas[,i] = pred_fun(model3.train,test_red,TRUE)
}

## Average across 100 iterations
model1_ave = round(apply(pred1_meas,1,mean),4)
model2_ave = round(apply(pred2_meas,1,mean),4)
model3_ave = round(apply(pred3_meas,1,mean),4)

model1_ave
```

## Prediciton Accuracy, k-fold and random sampling


## Average Prediction Accuracy: Comparing Data-Split Approaches 
```{r}
library(modelr)
library(purrr)

## Random Sampling: 20% Test Data for 100 iterations
cvrandom = crossv_mc(clean_data, n=100, test = 0.2)
cvrandom_red = crossv_mc(data_red, n=100, test = 0.2)

## K-fold Cross-Validation with K=10
cvkfold = crossv_kfold(clean_data, k=10)
cvkfold_red = crossv_kfold(data_red, k=10)

## Fit models for all combinations of k-1 folds 
models1kfold <- map(cvkfold$train, ~lm(cnt ~ ., data = .))
models2kfold <- map(cvkfold$train, ~lm(sqrt(cnt) ~ ., data = .))
models3kfold <- map(cvkfold_red$train, ~lm(sqrt(cnt) ~ ., data = .))

## Fit models for all random data splits
models1random <- map(cvrandom$train, ~lm(cnt ~ ., data = .))
models2random <- map(cvrandom$train, ~lm(sqrt(cnt) ~ ., data = .))
models3random <- map(cvrandom_red$train, ~lm(sqrt(cnt) ~ ., data = .))

## Obtain predictions across all models
get_pred_meas  <- function(model,test_data,transf=FALSE){
  data  <- as.data.frame(test_data)
  test  <- add_predictions(data, model)
  caret_model = mse_fun(test$pred,test$cnt) 
  mae_model =  mae_fun(test$pred,test$cnt)
  mape_model = mape_fun(test$pred,test$cnt) 
  pm_model  = pm_fun(test$pred,test$cnt)
  if(transf==TRUE){
    mse_model = mse_fun(test$pred^2,test$cnt)
    mae_model =  mae_fun(test$pred^2,test$cnt)
    mape_model = mape_fun(test$pred^2,test$cnt) 
    pm_model  = pm_fun(test$pred^2,test$cnt)
  }
  meas = c(mse_model,mae_model,mape_model,pm_model)
  return(meas)
}
pred1kfold = NULL
pred2kfold = NULL
pred3kfold = NULL
N = length(models1kfold)
for(i in 1:N){
  pred1kfold = cbind(pred1kfold,get_pred_meas(models1kfold[[i]],cvkfold$test[[i]]))
  pred2kfold = cbind(pred2kfold,get_pred_meas(models2kfold[[i]],cvkfold$test[[i]],TRUE))
  pred3kfold = cbind(pred3kfold,get_pred_meas(models3kfold[[i]],cvkfold_red$test[[i]],TRUE))
}
model1kfold_ave = round(apply(pred1kfold,1,mean),4)
model2kfold_ave = round(apply(pred2kfold,1,mean),4)
model3kfold_ave = round(apply(pred3kfold,1,mean),4)

pred1random = NULL
pred2random = NULL
pred3random = NULL
N = length(models1random)
for(i in 1:N){
  pred1random = cbind(pred1random,get_pred_meas(models1random[[i]],cvrandom$test[[i]]))
  pred2random = cbind(pred2random,get_pred_meas(models2random[[i]],cvrandom$test[[i]],TRUE))
  pred3random = cbind(pred3random,get_pred_meas(models3random[[i]],cvrandom_red$test[[i]],TRUE))
}
model1random_ave = round(apply(pred1random,1,mean),4)
model2random_ave = round(apply(pred2random,1,mean),4)
model3random_ave = round(apply(pred3random,1,mean),4)

```



```{r}
model1kfold_ave
model1random_ave
```

***
## P-values and Large Sample Size

### P-value Problem
```{r p-value iteration}
# Approach: Subsample 40% of the initial data sample & repeat 100 times
count = 1
n = nrow(clean_data)
B = 100
ncoef = dim(summary(model2)$coeff)[1]
pv_matrix = matrix(0,nrow = ncoef,ncol = B)

## Set percentage for the sub-sampled data
perc = 0.4 # 40% random sample of indices
while (count <= B) {
  subsample  = sample(n, floor(n*perc), replace=FALSE)
  # Extract the random subsample data
  subdata = clean_data[subsample,]
  # Fit the regression for each subsample
  submod = lm(sqrt(cnt)~.,data=subdata)
  # Save the p-values
  pv_matrix[,count] = summary(submod)$coeff[,4]
  # Increment to the next subsample
  count = count + 1
}

# Count pvalues smaller or equal to than 0.01 across the 100 (sub)models
alpha = 0.01
pv_significant = rowSums(pv_matrix <= alpha)

```


### Statistically Significant Coefficients
```{r extract significant}
# Which regression coefficients are statistically significant?
idx_scoef = which(pv_significant>=95)

# Show the p-values of the significant coefficients in model2
cbind(summary(model2)$coeff[idx_scoef,c(1,4)],
      Freq=pv_significant[idx_scoef])

# Plot the 100 p-values of the significant coefficients
matplot(pv_matrix[idx_scoef,],
     xlab="Regression Coefficient Index",
     ylab="P-values across 100 Samples",
     type="p",
     pch="o",
     col="blue")
```


### Coefficients Not Statsitically Significant
```{r}
## Which regression coefficients are not statistically significant?
idx_icoef = which(pv_significant<=85)

# Show the p-values of coefficients not stastically significant
# in model2
cbind(summary(model2)$coeff[idx_icoef,c(1,4)],
      Freq=pv_significant[idx_icoef])

# Plot the 100 p-values of the coefficients not statistically
# significant
matplot(pv_matrix[idx_icoef,],
     xlab="Regression Coefficient Index",
     ylab="P-values across 100 Samples",
     type="p",
     pch="o",
     col="blue")
```


### Implications of the tunning parameter Percent Subsample
```{r}
count = 1
perc = 0.2
pv_matrix.2 = matrix(0,nrow = ncoef,ncol = B)
while (count <= B) {
  subsample  = sample(n, floor(n*perc), replace=FALSE)
  # Extract the random subsample data
  subdata = clean_data[subsample,]
  # Fit the regression for each subsample
  submod = lm(sqrt(cnt)~.,data=subdata)
  # Save the p-values
  pv_matrix.2[,count] = summary(submod)$coeff[,4]
  # Increment to the next subsample
  count = count + 1
}
pv_significant.2 = rowSums(pv_matrix.2 < alpha)
idx_icoef.2 = which(pv_significant.2<85)
cbind(round(summary(model2)$coeff[idx_icoef.2,c(1,4)],3),
      Freq=pv_significant.2[idx_icoef.2])
```

