---
title: "Rental Bikes"
author: "NS"
date: "2023-03-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Packages

```{r}

library(tidyverse)
library(readr)
library(lubridate)
library(dplyr)
library(data.table)
library(leaps)
library(ggplot2)
library(dplyr)
library(MASS)
library(corrplot)
```

## Overview

**Objective:**
Create a poisson model to estimate the demand for bikes and test its accuracy and for overdispersion.

**Response Variables:**

• *Y* (Cnt): Total bikes rented by both casual & registered users together

**Qualitative Predicting Variables:**

*Season*: Season which the observation is made (1 = Winter, 2 = Spring, 3 = Summer, 4 = Fall)
*Yr*: Year on which the observation is made
*Mnth*: Month on which the observation is made
*Hr*: Day on which the observation is made (0 through 23)
*Holiday*: Indictor of a public holiday or not (1 = public holiday, 0 = not a public holiday)
*Weekday*: Day of week (0 through 6)
*Weathersit*: Weather condition (1 = Clear, Few clouds, Partly cloudy, Partly cloudy, 2 = Mist & Cloudy, Mist & Broken clouds, Mist & Few clouds, Mist, 3 = Snow, Rain, Thunderstorm & Scattered clouds, Ice Pallets & Fog)

**Quantitative Predicting Variables:**

*Temp*: Normalized temperature in Celsius
*Atemp*: Normalized feeling temperature in Celsius
*Hum*: Normalized humidity
*Windspeed*: Normalized wind speed

### Load Data

```{r load data}
data <- read.csv("Bikes.csv", header=T)
head(data)
```
## Preparing Data for Predicition

We have both qualititative and quantitive predicting variables, which means we have to change the qualitative to factors.
We are also removing the columns we don't need. This includes the record index, date, count of casual users, and count of registered users

```{r prepare data, convert}

# Remove irrelevant columns
clean_data = data[-c(1,2,9,15,16)]
# Convert the numerical categorical variables to predictors
clean_data$season = as.factor(clean_data$season)
clean_data$yr = as.factor(clean_data$yr)
clean_data$mnth = as.factor(clean_data$mnth)
clean_data$hr = as.factor(clean_data$hr)
clean_data$holiday = as.factor(clean_data$holiday)
clean_data$weekday = as.factor(clean_data$weekday)
clean_data$weathersit = as.factor(clean_data$weathersit)

```

Next set to training nad testing
```{r training and testing sets}
# clean_data has categorical variables converted to factors , not necessary for doing separately on train and test
sample_size = floor(0.8*nrow(clean_data))
picked = sample(seq_len(nrow(clean_data)), size=sample_size)
train = clean_data[picked,]
test = clean_data[-picked,]
```

## Poisson Regression Analysis

We do a poisson regression to the new clean set.
We do poisson becasue the constant variance assumptin is violated when using MLR
```{r Poisson}
model1 = glm(cnt~., data=clean_data, family = "poisson")
summary(model1)
```

There are a lot of variables here, which can lead to inflated statistical significance.

## Poisson GOF
```{r Poisson residual analysis}
resids1 <- resid(model1, type='deviance')
hist(resids1, nclass=20, col="blue", border="gold" , main = "histogram of residuals")
qqnorm(resids1, col="blue")
qqline(resids1, col = 'red')

```
The normality assumption looks ok from the histogram, but the qqnorm seems to be heavy-tailed.

```{r GOF residuals}
with(model1, cbind(res.deviance = deviance, df = df.residual, p = pchisq(deviance, df.residual, lower.tail=FALSE)))
```

## Prediction Accuracy Measures

This is simliar to what we used for MLR in teh previous module. We use the same set of functions, and all together aggregated.

```{r Prediction Accuracy Measures and Aggregated}
# Mean Squared Prediction Error (MSPE)
mse_fun <- function(pred,dat){mean((pred-dat)^2)}
# Mean Absolute Prediction Error (MAE)
mae_fun <- function(pred,dat){mean(abs(pred-dat))}
# Mean Absolute Percentage Error (MAPE)
mape_fun <- function(pred,dat){mean(abs(pred-dat)/abs(dat))}
# Precision Measure (PM)
pm_fun <- function(pred,dat){sum((pred-dat)^2)/sum((dat-mean(dat))^2)}
## Aggregate Prediction Function
pred_fun <- function(model,test){ 
  pred = predict(model, test, type="response") 
  test.pred = pred 
  mse_model = mse_fun(test.pred,test$cnt)
  mae_model = mae_fun(test.pred,test$cnt)
  mape_model = mape_fun(test.pred,test$cnt)
  pm_model = pm_fun(test.pred,test$cnt)
  pred_meas = c(mse_model,mae_model, mape_model, pm_model)
  return(pred_meas)
}
```

## prediction accuracy: Poisson with Test/Train

We can measure the accuracy once or 100 times. 

```{r Accuracy measure for Poisson 1x}
## Accuracy measures for 1 iteration (Poisson Regression)
set.seed(0)
sample_size = floor(0.8*nrow(clean_data))
picked = sample(seq_len(nrow(clean_data)), size=sample_size)
train = clean_data[picked,];
test = clean_data[-picked,]
model1.train = glm(cnt~.,data=train,family="poisson")
pred_fun(model1.train,test)

```
These four are MSPE, MAE, MAPE, and PM. PM is t0.245, meaning about 25% of model is explained.


```{r Accuracy measure for Poisson 100x}
## Accuracy measures for 100 iteration (Poisson Regression)
set.seed(0)
pred1_meas = matrix(0,4,100)
for(i in 1:100){
  sample_size = floor(0.8*nrow(clean_data))
  picked = sample(seq_len(nrow(clean_data)), size=sample_size)
  train = clean_data[picked,]; test = clean_data[-picked,]
  model1.train = glm(cnt~.,data=train,family="poisson")
  pred1_meas[,i] = pred_fun(model1.train,test)
}
model1_ave = round(apply(pred1_meas,1,mean),4)
model1_ave

```

For 100x, it doesn't seem to be all that much better

## P-value and inflated significance, subsampling

instead of all data, we do 20% with 100 repetitions and apply the poisson regression
**Tunning Parameter**: percent sub-sample

```{r tunning parameter, sub-sample}
## Approach: Subsample 20% of the initial data sample & repeat 100 times
count = 1
n = nrow(clean_data)
B = 100 #repetitions
ncoef = dim(summary(model1)$coeff)[1] #no of coefficients
pv_matrix = matrix(0,nrow = ncoef,ncol = B)
while (count <= B){
  subsample = sample(n, floor(n*0.2), replace=FALSE) #sample 20%
  subdata = clean_data[subsample,] #sample that from original set
  # Fit the poisson regression for each subsample
  submod = glm(cnt~.,data=subdata,family="poisson")
  ## Count pvalues smaller than 0.01 across the 100 (sub)models
  pv_matrix[,count] = summary(submod)$coeff[,4]
  count = count + 1
}
alpha = 0.01
pv_significant = rowSums(pv_matrix < alpha)
```

## Identifying Statistical Significance
The above gave us 100 values and we are testing them against a sig level of95% larger than sig level.

```{r find reg coeff that are significant}
# Identify variables which have p-values less<alpha on more than 95% iterations
idx_scoef = which(pv_significant>=95)
length(idx_scoef)
```
This means 44/51 p-values are small across the sub-samples

We can also plot this
```{r sub-sampling plot}
matplot(pv_matrix[idx_scoef,],
        xlab="Regression Coefficient Index",
        ylab="P-values across 100 Samples",
        type="p",
        pch="o",
        col="blue")

```

## Lack Statistical Significance

We want to ind which coeficients are not statistically signficant, which would mean a pv less than 85
```{r lack stat sig}
idx_icoef = which(pv_significant<85)
# Show the p-values of the non-significant coefficients in model2
cbind(summary(model1)$coeff[idx_icoef,c(1,4)],
      Freq=pv_significant[idx_icoef])
# Plot the 100 p-values of the non-significant coefficients
matplot(pv_matrix[idx_icoef,],
        xlab="Regression Coefficient Index",
        ylab="P-values across 100 Samples",
        type="p",
        pch="o",
        col="blue")
```
We can see from the table which values were not sinificant because 80 of p-values smalle rthan significance
the distribution of p-values is approx uniform

## Overdispersion

Like the other models we can have overdispersion. We can find this by plotting the log of fitted values against log of squared differences.
With Overdispersion, we can try to correct with quasi poisson or negative binomial.

```{r overdispersion plot}
# Overdispersion (a probable cause of inflated significance)
plot(log(fitted(model1)),log((data$cnt-fitted(model1))^2),xlab=expression(hat(mu)),ylab=expression((y-hat(mu))^2),pch=20,col="blue")
abline(0,1) ## 'variance = mean' line

```

We can see that the overdispersion occurs above the line here, which means that most observations, variance is higher than the mean.

### Overdispersion parameter
```{r overdispersion parameter}
#overdispersion parameter

dp = sum(residuals(model1,type ="pearson")^2)/model1$df.residual
dp
```

It's above 32, which is WAY more than the threshold of 2 or 4.


```{r}
# see how the coefficients are affected owing to overdispersion
cbind(original_estimates=summary(model1)$coeff[,4],dispersion_estimates=summary(model1,dispersion=dp)$coeff[,4])
```


```{r}
# effect of overdispersion on model coefficients 

length(which(summary(model1,dispersion=dp)$coeff[,3]<0.025))
length(which(summary(model1)$coeff[,3]<0.025))
```
